\todo{How's this title?}A Survey of Web Testing Frameworks and Workflow for Testing JavaScript Applications

### 2014-06-01 Paper/paper.tex
Cleaned up a bit, added back in TODO notes.

% Two Column Format
\documentclass[12pt]{ucthesis}
%this allows us to specify sections to be single or multi column so that things like title page and table of contents are single column
\usepackage{listings}
\usepackage{placeins}
\usepackage{caption}
\usepackage[disable]{todonotes}
\usepackage{url}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[letterpaper]{geometry}
\usepackage[overload]{textcase}
\usepackage{verbatim}

@@

\title{A Survey of Web Testing Frameworks and Workflow for Testing JavaScript Applications} %\vfill gives us the black space at the top of the page
\author{Taggart Ashby}
\degreemonth{June} \degreeyear{2014} \degree{Master of Science}
\defensemonth{June} \defenseyear{2014}
\numberofmembers{3} \chair{Michael Haungs, Ph.D.} \othermemberA{Franz Kurfess, Ph.D.} \othermemberB{David Janzen, Ph.D.} \field{Computer Science} \campus{San Luis Obispo}
\copyrightyears{seven}

@@

%%%%%%%%%%%%%%%%
%%% Abstract 
%%% VERY General overview of problem and solution in paper
%%%%%%%%%%%%%%%%
\begin{abstract}
\todo{Make this better}
Web applications are quickly replacing standalone applications. As with any type of application, these web applications need to be tested to ensure proper functionality and reliability. There have been substantial efforts to create tools to help with the testing of web applications but there is, as of yet, no standard set of tools or recommended workflow to insure speed of development and strength of application.

@@

\section{What Has Been Done Already?}
Web developers are not clueless, they understand that there is a need for testing of web applications. Unfortunately, web development is still a relatively new domain and developers have not come together to work on the problem of testing but rather have come up with their own proprietary solutions in isolation or merely on a per-project basis. Some of these tools have gained traction and become larger, more general, tools.

@@

\todo{What goes here?}
\todo{Info info info}

\section{Problems Still Faced/Why It's So Hard}
% - A lot of products to test against (browsers)
Testing web applications is difficult. The web offers an impressive set of features, but with those features come some draw-backs. Two of the most frustrating drawbacks for developers are the number of platforms and the web's asynchronous nature.

@@

%\todo{THIS NEEDS TO MOVE OR CHANGE SUBSECTION TO `WEB APPLICATIONS'}
\subsubsection{Node.JS}
As a quick aside, it will be useful to have information about Node.JS as it is used in the sample application and the evaluation application. Node.JS, often called simply Node, is a JavaScript API for creating and running servers. ``Node's goal is to provide an easy way to build scalable network programs.'' \cite{Node} Node was created in 2009 and is sponsored by Joyent. Node does not use thread-based networking, instead opting for a single-threaded event loop and non-blocking I/O. Node allows developers to create and control web servers without the need for external software, such as Apache, commonly found in other sites.

@@

\todo{TIE THIS CLIENT/SERVER STUFF into Web Applications, be specific, not generic}
\section{Client/Server Architecture}
Web applications have a user-facing client-side and a ``rear-facing'' \todo{IS THIS THE RIGHT TERM...} server-side.

@@

\subsection{What is the Client-side Portion?}
First, we'll address what a client is. The simple answer is: The User. The User's phone, computer, tablet, or other device communicating with a given website or application is the client. More specifically, the browser or service being using to communicate with the web is the client.

@@

\subsection{What is the Server-side Portion?}
A server is the entity, be it machine, person, or cloud service, that stores the information about a given web page or application. The server stores all assets: HTML pages, images, videos, databases, etc, and provides them upon request to the client. The server also handles all of the processing for interactivity, accounts, and general number-crunching.

@@

\section{What is Testing?}
IEEE defines software testing as ``the dynamic verification of the behavior of a program on a finite set of test cases, suitably selected from the usually infinite executions domain, against the expected behavior.'' \cite{TestingDefinition} Put a bit more simply, software testing is the process of making sure the program does what it is supposed to.

@@

\section{ToDo App}
In order to evaluate the various frameworks that we found online, we needed a small application with some functionality, but without the inherent complexity that was going to come with our final evaluation application. We chose a ToDo application made with Express.js, Node.js, and MongoDB. \cite{ToDoAppHomePage} Our choice of this application comes from its limited functionality and the ease of understanding the interactions that take place within the application, i.e. adding tasks to a ToDo list, making tasks as completed, and viewing completed tasks. Here is a brief preview of the application:

@@

%%%%%%%%%%%%%%%%%%%%%%
%%% Testing Frameworks
%%%   Talk about the various frameworks
%%%%%%%%%%%%%%%%%%%%%%
\section{Testing Frameworks}

% NOT SURE I NEED THIS SECTION AT ALL
% \section{Assumptions and Requirements}
% In order to more thoroughly vet the many testing frameworks we came across, we had a number of assumptions about the projects we'd be looking at and requirements for those framework. 
%- Node.js server because it's JS, so it's testable.
%- Something to do unit testing
%- Something to do continuous integration

@@

%- Node.js
%- Browser based
%- JavaScript
%- Database
%- Open-source

\subsection{Tool Requirements}
\todo{Move this section elsewhere}
Our requirements for the selection of testing tools was relatively simple. In addition to meeting the requirements of our ``average web application'', the only other requirement was that the tool was free. What ``free'' means is that the developer would have access to at least the minimum functionality needed to use the tool on \emph{one} application that is \emph{open-source}. What we quickly found is that web testing tools love open-source projects but are far less willing to provide free functionality to private or closed-source projects.

Occasionally the cost of tools will be discussed but in general we did not review tools that only offered paid options.
%- Free

@@

\subsubsection{Assertion Library of Choice: Chai.js \cite{Chaijs}}
For the assertion library we chose Chai.js.  It offers the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and a more general Assert API. The Expect API provides assertions in the form of \lstinline{expect(myVar).to.equal(10);} or \lstinline{expect(myArr).to.have.length(10);}. The Should API has assertions of the form \lstinline{myVar.should.be.a('number');} or \lstinline{myArr.should.have.length(10);}. Finally, the Assert API has the more tried and true \lstinline{assert.equal(myVar, 10);} and \lstinline{assert.lengthOf(myArr, 10);}. The various APIs combined with Mocha's flexibility led to a nice developer experience that can appeal to a variety of different developers.

\subsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the \lstinline{it} functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png}

\captionof{figure}{Each individual spec is further broken down into the assertions within it.}

\end{Figure}
An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions are expected to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let the developer know fewer assertions were called than expected.

@@

The difference between PhantomJS and ZombieJS, which I talk about in the next section, is akin to the difference between Selenium and Nightwatch. PhantomJS is heavier duty and requires an additional set of testing tools whereas ZombieJS is lightweight and has assertions built-in.
% Too much.
%- Really great though, could be useful if needed.
%- Slow
%- FULL DOM/Browser emulation

@@

%- Lightweight
%- Headless
%- NOT Full DOM emulation

@@

\section{Misc Frameworks}
In addition to the Unit Testing, Continuous Integration, and Browser Testing frameworks outlined above, we came across a number of incredibly helpful frameworks that did not quite fit into any of the above categories.

@@

\section{Framework Decision Matrices}

\FloatBarrier
\begin{table}[h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Framework & Async Support & Free & Separate Server & Assertion Styles & Test Interfaces & Reporters \\ \hline
Jasmine   & X             & X    &                 &                  &                 &           \\ \hline
Mocha     & X             & X    &                 & X*               & X               & X         \\ \hline
Qunit     & X             & X    &                 &                  &                 &           \\ \hline
Intern    & X             & X    & X               & X                & X               & X         \\ \hline
\end{tabular}
}
\captionof{table}{Decision Matrix for Unit Testing Frameworks}
\end{table}
\FloatBarrier
\footnotesize

@@

\section{Workflow}
In addition to just a list of tools and installation instructions, we'd like to propose a way to use those tools that is efficient and nearly autonomous. Our goal is to bring testing tools into use with as little resistance as possible. With a bit of effort at the outset, testing can become a crucial and enjoyable part of development.

@@

All of the unit testing code for the PolyXpress Author is available in the appendix.

@@

\subsubsection{Headless Browsers}
The notion of headless browsers is extremely interesting and worth of exploration. PhantomJS appeared to be quite promising for a full DOM implementation or ZombieJS for more simple DOM emulation. A side-by-side comparison of tests running through a full-fledged browser and tests running through a headless browser would be interesting and enlightening as to whether or not headless browsers could have a real place in a testing framework.

>>>>

% Two Column Format
\documentclass[12pt]{ucthesis}
%this allows us to specify sections to be single or multi column so that things like title page and table of contents are single column
\usepackage{listings}
\usepackage{placeins}
\usepackage{caption}
\usepackage{todonotes}
\usepackage{url}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[letterpaper]{geometry}
\usepackage[overload]{textcase}
\usepackage{verbatim}

@@

\title{\todo{How's this title?}A Survey of Web Testing Frameworks and Workflow for Testing JavaScript Applications} %\vfill gives us the black space at the top of the page
\author{Taggart Ashby}
\degreemonth{June} \degreeyear{2014} \degree{Master of Science}
\defensemonth{June} \defenseyear{2014}
\numberofmembers{3} \chair{Michael Haungs, Ph.D.} \othermemberA{Franz Kurfess, Ph.D.} \othermemberB{David Janzen, Ph.D.} \field{Computer Science} \campus{San Luis Obispo}
\copyrightyears{seven}

@@

%%%%%%%%%%%%%%%%
%%% Abstract 
%%% VERY General overview of problem and solution in paper
%%%%%%%%%%%%%%%%
\begin{abstract}
Web applications are quickly replacing standalone applications. As with any type of application, these web applications need to be tested to ensure proper functionality and reliability. There have been substantial efforts to create tools to help with the testing of web applications but there is, as of yet, no standard set of tools or recommended workflow to insure speed of development and strength of application.

@@

\section{What Already Exists}
Web developers are not clueless, they understand that there is a need for testing of web applications. Unfortunately, web development is still a relatively new domain and developers have not come together to work on the problem of testing but rather have come up with their own proprietary solutions in isolation or merely on a per-project basis. Some of these tools have gained traction and become larger, more general, tools.

@@

\section{Problems Still Faced}
% - A lot of products to test against (browsers)
Testing web applications is difficult. The web offers an impressive set of features, but with those features come some draw-backs. Two of the most frustrating drawbacks for developers are the number of platforms and the web's asynchronous nature.

@@

\subsubsection{Node.JS}
As a quick aside, it will be useful to have information about Node.JS as it is used in the sample application and the evaluation application. Node.JS, often called simply Node, is a JavaScript API for creating and running servers. ``Node's goal is to provide an easy way to build scalable network programs.'' \cite{Node} Node was created in 2009 and is sponsored by Joyent. Node does not use thread-based networking, instead opting for a single-threaded event loop and non-blocking I/O. Node allows developers to create and control web servers without the need for external software, such as Apache, commonly found in other sites.

@@

\section{Client/Server Architecture}
Web applications have a user-facing client-side and a ``rear-facing'' \todo{Is this the right term?} server-side.

@@

\subsection{Client-side}
First, we'll address what a client is. The simple answer is: The User. The User's phone, computer, tablet, or other device communicating with a given website or application is the client. More specifically, the browser or service being using to communicate with the web is the client.

@@

\subsection{Server-side}
A server is the entity, be it machine, person, or cloud service, that stores the information about a given web page or application. The server stores all assets: HTML pages, images, videos, databases, etc, and provides them upon request to the client. The server also handles all of the processing for interactivity, accounts, and general number-crunching.

@@

\section{Defining Testing}
IEEE defines software testing as ``the dynamic verification of the behavior of a program on a finite set of test cases, suitably selected from the usually infinite executions domain, against the expected behavior.'' \cite{TestingDefinition} Put a bit more simply, software testing is the process of making sure the program does what it is supposed to.

@@

\section{ToDo Application}
In order to evaluate the various frameworks that we found online, we needed a small application with some functionality, but without the inherent complexity that was going to come with our final evaluation application. We chose a ToDo application made with Express.js, Node.js, and MongoDB. \cite{ToDoAppHomePage} Our choice of this application comes from its limited functionality and the ease of understanding the interactions that take place within the application, i.e. adding tasks to a ToDo list, making tasks as completed, and viewing completed tasks. Here is a brief preview of the application:

@@

@@

\subsection{Tool Requirements}
Our requirements for the selection of testing tools was relatively simple. In addition to meeting the requirements of our ``average web application'', the only other requirement was that the tool was free. What ``free'' means is that the developer would have access to at least the minimum functionality needed to use the tool on \emph{one} application that is \emph{open-source}. What we quickly found is that web testing tools love open-source projects but are far less willing to provide free functionality to private or closed-source projects.

Occasionally the cost of tools will be discussed but in general we did not review tools that only offered paid options.

@@

\subsubsection{Chai.js \cite{Chaijs}}
For the assertion library we chose Chai.js.  It offers the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and a more general Assert API. The Expect API provides assertions in the form of \lstinline{expect(myVar).to.equal(10);} or \lstinline{expect(myArr).to.have.length(10);}. The Should API has assertions of the form \lstinline{myVar.should.be.a('number');} or \lstinline{myArr.should.have.length(10);}. Finally, the Assert API has the more tried and true \lstinline{assert.equal(myVar, 10);} and \lstinline{assert.lengthOf(myArr, 10);}. The various APIs combined with Mocha's flexibility led to a nice developer experience that can appeal to a variety of different developers.

\subsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the \lstinline{it} functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png}
  \captionof{figure}{Each individual spec is further broken down into the assertions within it.}
\end{Figure}

An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions are expected to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let the developer know fewer assertions were called than expected.

@@

The difference between PhantomJS and ZombieJS, which I talk about in the next section, is akin to the difference between Selenium and Nightwatch. PhantomJS is heavier duty and requires an additional set of testing tools whereas ZombieJS is lightweight and has assertions built-in.

@@

@@

\section{Miscellaneous Frameworks}
In addition to the Unit Testing, Continuous Integration, and Browser Testing frameworks outlined above, we came across a number of incredibly helpful frameworks that did not quite fit into any of the above categories.

@@

\section{Framework Decision Matrices}
\todo{These are terrible... I have no idea what to put in these tables for columns.}
\FloatBarrier
\begin{table}[h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Framework & Async Support & Free & Separate Server & Assertion Styles & Test Interfaces & Reporters \\ \hline
Jasmine   & X             & X    &                 &                  &                 &           \\ \hline
Mocha     & X             & X    &                 & X*               & X               & X         \\ \hline
Qunit     & X             & X    &                 &                  &                 &           \\ \hline
Intern    & X             & X    & X               & X                & X               & X         \\ \hline
\end{tabular}
}
\captionof{table}{Decision Matrix for Unit Testing Frameworks}
\end{table}
\FloatBarrier
\footnotesize

@@

\section{Suggested Workflow}
In addition to just a list of tools and installation instructions, we'd like to propose a way to use those tools that is efficient and nearly autonomous. Our goal is to bring testing tools into use with as little resistance as possible. With a bit of effort at the outset, testing can become a crucial and enjoyable part of development.

@@

All of the unit testing code for the PolyXpress Author is available in the appendix. The mocks are at the top of each test file, though the Model does not have any mocks.

@@

\subsubsection{Headless Browsers}
The notion of headless browsers is extremely interesting and worthy of exploration. PhantomJS appeared to be quite promising for a full DOM implementation or ZombieJS for more simple DOM emulation. A side-by-side comparison of tests running through a full-fledged browser and tests running through a headless browser would be interesting and enlightening as to whether or not headless browsers could have a real place in a testing framework.
